---
title: "Lab 1"
author: "Danial Shaikh"
format: 
  html:
    self-contained: true
execute:
  echo: true
  error: true
  
editor: visual
---

## Cars Dataset

*This dataset contains information about \_\_the speed of cars and the distances taken to stop\_\_. The data was collected in \_\_the 1920s\_\_.*

*It has \_50 observation\_\_ rows and \_\_2\_\_ variables.*

```{r}

summary(cars)

```

## Pressure Data Set

```{r}
plot(pressure, xlab="Temperature(Degrees Celsius)",ylab="Pressure of mercury(mm)")
```

## Cars Data Visualizations

```{r}
dist<-cars[,2]
plot(dist, xlab='Stopping distance (ft)')
plot (cars,xlab= 'Car Speed (mph)', ylab='Stopping distance (ft)')
```

The first chart shows the distributions of stopping distances

the second chart shows stopping distances in respect to car speeds.

there is a correlation between higher car speeds and longer stopping distances.

## Improvements I would like to make

-   I would like to change the colors of the dots to make them more viable

-   I would like to add a trend line to better see correlation.
