---
title: "Lab4"
author: "Danial Shaikh"
format:
  html:
    self-contained: true
    code-tools: true
    code-fold: true
execute: 
  echo: true
  error: true
---

```{r}
library(tidyverse)
library(dplyr)
library(here)

```

```{r}

# Code for question 2.0 
avocado <- read_csv(here("Week4","avocado.csv"))
avocado
```

Question 2.1

The dataset contains information about avocado prices and sales in the United States from 2015 to 2018. The information includes the following variables:

-   Date: the date of the observation

-   AveragePrice: average price of a single avocado

-   Total Volume: total number of avocados sold

-   4046: total number of avocados with PLU 4046 sold

-   4225: total number of avocados with PLU 4225 sold

-   4770: total number of avocados with PLU 4770 sold

-   Total Bags: total number of bags of avocados sold

-   Small Bags: total number of small bags of avocados sold

-   Large Bags: total number of large bags of avocados sold

-   XLarge Bags: total number of extra large bags of avocados sold

-   type: conventional or organic

-   year: year of the observation

-   region: geographical region in the US where the observation was made.

```{r}
avocado_clean <- avocado
length(unique(avocado_clean$region))


```

```{r}
avocado_clean |>
  rename(
    Small = `4046`, Large = `4225`,xtraLarge = `4770`)
major_regions <- c("Plains", "Midsouth", "West", "SouthCentral", "Northeast", "Southeast")
  
major_regions <- avocado_clean |>
  filter(region %in% major_regions) |>
  rename(
    Small = `4046`, Large = `4225`,xtraLarge = `4770`)


metro_regions <- avocado_clean |>
  filter(!(region %in% c(major_regions, "TotalUS")))|>
  rename(
    Small = `4046`, Large = `4225`,xtraLarge = `4770`)

major_regions
```

\

**3.** Which major region sold the most organic, small Hass avocados in 2017?\

```{r}
major_regions |> 
  filter(type == "organic", year == 2017) |>
  group_by(region) |>
  summarize(tot_small_bags = sum(Small)) |>
  arrange(desc(tot_small_bags)) |>
  slice(1)
```

**4.** Use the `separate()` function to split the `Date` variable into year, month, and day. In which month is the highest volume of avocado sales?

\

```{r}
avocado_clean |>
  separate(Date, into = c("Year", "Month", "Day"), sep = "-") |>
  group_by(Month) |>
  summarize(tot_sales = sum(`Total Volume`)) |>
  arrange(desc(tot_sales)) |>
  slice(1)
```

**5.** Which metro area regions sold the most total avocados? Plot side-by-side boxplots of total volume for only the five regions with the highest averages for the `Total Volume` variable.

**\
**

```{r}
mostAvo5 <- metro_regions |>
    group_by(region) |>
    summarise(average_total_volume = mean(`Total Volume`)) |>
    arrange(desc(average_total_volume)) |>
    head(5) |>
    pull(region)

ggplot(data = avocado_clean |> filter(region %in% top5regions),
       aes(x = region, y = `Total Volume`, fill = region)) +
  geom_boxplot()

```

**6.** In which regions is the price of organic versus conventional avocados most different? Support your answer with a few summary statistics and a visualization.

```{r}
cali <- avocado_clean |> 
  filter(region %in% c("LosAngeles", "SanDiego", "Sacramento", "SanFrancisco"))

ggplot(data = cali, aes(x = region, y = AveragePrice, fill = type)) + 
  geom_boxplot() +
  ggtitle("Prices of Avocados in different California Regions")

cali |> 
  group_by(region, type) |>
  summarise(average_price = mean(AveragePrice)) |>
  pivot_wider(names_from = type, values_from = average_price, values_fill = 0) |>
  mutate(
    price_difference = organic - conventional
  )
```

**\
**

```{r}
cali
```

**7.** The following plot shows, for all four California regions, the proportion of the mean Hass avocado sales that are small, large, or extra large; conventional vs. organic. Recreate the plot.

```{r}
cali_grouped <- group_by(cali_avo, region, type)
cali_summarized <- summarize_at(cali_grouped, vars(Size_Small, Size_Large, Size_XL), funs(mean))
cali_gathered <- gather(cali_summarized, key = "Size", value = "Num.Sold", -region, -type)
cali_mutated <- mutate(cali_gathered, 
                       Size = factor(Size, 
                                     levels = c("Size_Small", "Size_Large", "Size_XL"),
                                     labels = c("Small", "Large", "Extra Large"))
                      )
cali_plotted <- ggplot(cali_mutated, aes(x = region, y = Num.Sold, fill = Size)) +
                geom_col(position = "fill") + 
                facet_wrap(~type)
cali_plotted

```

```{r}

```
